### 2025-12-19 Paper/paper_2025.tex

   To create a measure of board/executive interlock, I compile data on nonprofit hospital board of directors and executives in years 2017-2022 from publicly available Tax Form 990s. All sufficiently large nonprofits must file these forms with the Internal Revenue Service (IRS) each year. They contain a section in which nonprofits declare their board, executives, and highest compensated employees. This paper focuses only on the board of director identities. I limit to firms that fill out Schedule H of the tax form, an indicator that the firm operates a hospital. I extract all names and positions of individuals listed in the section titled ``Officers, Directors, Trustees, Key Employees, and Five Highest Compensated Employees". I also record information on the hospital that filed the tax form: the Employee Identification Number (EIN), organization name and location, and other characteristics. The initial extraction yields 2,096 nonprofit hospitals.

@@

    I record self-reported information about the individuals in the data, including whether they have a title for medical doctor, nurse, or having a degree in health administration. To clean the names, I remove any titles or credentials, and I convert common nicknames to their longer versions for consistency. I also combine names within the same firm that have either slight misspelling differences or, if middle names are only sometimes listed, they have at least 2 names in common. For example, if John Lee Matthew is a board member of hospital A in 2017, and John Matthew is a board member of hospital A in 2018, I assume these names represent the same person. Finally, I use the gender package in R to predict each individual's gender using historical data (\cite{gender}).

    I present summary statistics of characteristics of the people in the data in Table \ref{tab:board_people}. There are just under 37,000 people in the data, 63\% of which are board members and the remaining 37\% are executives. Board members are slightly moe likely to be medical doctors than executives, but the fraction of doctor board members is still relatively low at 18\%. Nurses are consistent across board and executives, but only 3\% of the sample are nurses. A negligible number of people claim to have a health administration degree. Executives are slightly more likely to be female according to the name algorithms, with 38\% of executives estimated to be female and 32\% of board members estimated to be female. I also include average team composition of hospital boards and executive teams, where these trends largely hold at the hospital level. On average, hospital boards contain 7 members and executive teams contain 10 members.

@@

    I create a list of overlapping board member candidates by identifying any names that occur on multiple hospital boards in the same year. However, it is very common for hospitals within a system to largely share the same board. This is not the type of variation I seek to investigate in this paper, since hospitals within systems are already heavily affiliated. Therefore, I define the relevant type of overlap as someone being on the board of multiple hospitals that are otherwise \textit{unaffiliated}. Hospitals within systems are still eligible to have overlap with other hospitals that do not belong to the same system. I exclude hospitals that gain or lose overlapping board members several times in the sample period.

    In the hospital merger literature, mergers have been found to be more significant when the two hospitals involved are located geographically close (\cite{cooper2019price}). Therefore, I define two types of overlap: overlap with an otherwise non-affiliated hospital in the same market, or overlap with an otherwise unaffiliated hospital in a different market. I use the market definitions developed by Dartmouth Atlas, Hospital Referral Regions (HRRs), which are meant to capture geographic regions of the average patient's choice set of hospitals.

@@

    In Figure \ref{fig:connected_percent} I show the prevalence of overlapping board members or executives over time. In 2017, approximately 5\% of hospitals have an overlapping board members or executive with another hospital in the same market that is not formally affiliated otherwise. That is, these hospital are not under common ownership. This percentage increases over time, with 10\% of hospitals sharing a key leader in their market in 2022. I also graph the percent of HRR markets where hospitals within that market share board members or executives. Approximately 15\% of HRRs are potentially affected by this type of affiliation. Sentence here about putting this into perspective relative to hospital consolidation as a whole.

@@

    Next, I present the geographic distribution of these pairs in Figure \ref{fig:connected_maps}. The blue lines represent hospitals that share a common board member, and the red dots are all other hospitals. Connected hospitals are distributed fairly evenly across the US, apart from relatively few connections on the west coast relative to the population. Over time, the connections remain fairly consistent geographically.

    \begin{figure}[ht!]
        \centering
        \caption{Geographic Distribution of Board/Executive Interlock}
        \includegraphics[width=\textwidth]{Objects/connected_maps.pdf}
        \label{fig:connected_maps}
    \end{figure}

    Hospitals in the sample are either always connected, never connected, or become connected at some point in time. I show characteristics of hospitals in these categories in Table \ref{tab:hosp_group_stats}.

@@

    I now present characteristics of pairs of hospitals that are connected by board members or executives in Table \ref{hospital_pair_types}. In the first column, I show pair characteristics of hospitals connected within the same HRR. In the second column, I present characteristics of hospitals connected to otherwise unaffiliated hospitals located in a different HRR. Therefore, all hospital pairs in the table are connected, but only the first column are connected to hospitals that are potentially direct competitors. Each hospital in a pair is general or specialty (I only consider adult hospitals for these categories), children or adult, and independent or belonging to a system, all characteristics in the AHA data. Additionally, the median bed size of all hospitals in the sample is 101, and I define hospitals as being either small or large depending on whether they fall below or above this median. There are typically 1.5 connections in the same HRR and almost 6 connections in different HRRs.

@@

    Most pairs, 76-78\% are two general hospitals. Thus, for the majority of hospitals that have board/executive interlock in the same HRR, the hospital they share a leader with is a direct competitor. I also include characteristics of system versus independent ownership. For there is a nearly even spread of the type of ownership of the pair. 42\% of these pairs are made up of one independent hospital and one system-owned hospital, while 37\% are made up of two independent hospitals and the rest are made up of two hospitals that are both system owned but belong to two different systems. These figures are slightly different for connections in different HRRs, where a larger portion of connections, 45\%, are made up of two system-owned hospitals of two different systems.

    In both same and different HRR connected pairs, 50\% of the connections are such that one hospital is small in bed size and one is large in bed size. The average bed difference is over 200, indicating that the size difference of the average hospital connection is drastic. what's the implication of this? Finally, connected hospitals in the same HRR are approximately 62 kilometers, or 38 miles, away from the other.

    \import{Objects}{hosp_outcomes_table.tex}

>>>>

   To create a measure of board/executive overlap, I compile data on nonprofit hospital board of directors and executives in years 2013-2023 from publicly available Tax Form 990s. All sufficiently large nonprofits must file these forms with the Internal Revenue Service (IRS) each year. They contain a section in which nonprofits declare their board, executives, and highest compensated employees. This paper focuses only on board members and executives. I limit to firms that fill out Schedule H of the tax form, an indicator that the firm operates a hospital. I extract all names and positions of individuals listed in the section titled ``Officers, Directors, Trustees, Key Employees, and Five Highest Compensated Employees". I also record information on the hospital entity: the Employee Identification Number (EIN), organization name and location, and other characteristics. The initial extraction yields 2,096 nonprofit hospitals.

@@

    I maintain information about the individuals in the data, including whether they have a title for medical doctor, nurse, or having a degree in health administration. To clean the names, I remove any titles or credentials, and I convert common nicknames to their longer versions for consistency. I also combine names within the same firm that have either slight misspelling differences or, if middle names are only sometimes listed, they have at least 2 names in common. For example, if John Lee Matthew is a board member of hospital A in 2017, and John Matthew is a board member of hospital A in 2018, I assume these names represent the same person. Finally, I use the gender package in R to predict each individual's gender using historical data (\cite{gender}).

    I present summary statistics of characteristics of the people in the data in Table \ref{tab:board_people}. There are just under 52,000 people in the data at some point over the course of 10 years, 63\% of which are board members and the remaining 37\% are executives. Board members are slightly more likely to be medical doctors than executives, but the fraction of doctor board members is still relatively low at 18\%. Nurses are consistent across board and executives, but only 3\% of the sample are nurses. A negligible number of people claim to have a health administration degree. Executives are slightly more likely to be female according to the name algorithms, with 38\% of executives estimated to be female and 30\% of board members estimated to be female. I also include average team composition of hospital boards and executive teams, where these trends largely hold at the hospital level. On average, hospital boards contain 19 members and executive teams contain 14 members.

@@

    I create a list of overlapping board member or executive candidates by identifying any names associated with multiple hospital in the same year. However, it is very common for hospitals within a system to largely share the same board members and/or executives. This is not the type of variation I seek to investigate in this paper, since hospitals within systems are already heavily affiliated. Therefore, I define the relevant type of overlap as someone being on the leadership team of multiple hospitals, where the hospitals are otherwise unaffiliated by system or network. Hospitals within systems are still eligible to have overlap with other hospitals that do not belong to the same system. I exclude hospitals that gain or lose overlapping board members several times in the sample period.

    In the hospital merger literature, mergers have been found to be more significant when the two hospitals involved are located geographically close (\cite{cooper2019price}). Therefore, I define two types of overlap: overlap with an otherwise non-affiliated hospital in the same market, or overlap with an otherwise unaffiliated hospital in a different market. I use the market definitions developed by Dartmouth Atlas, Hospital Referral Regions (HRRs), which are meant to capture geographic regions of the average patient's choice set of hospitals. From these overlap definitions, I characterize hospitals by the first and last year they are connected in the same HRR, and the first and last year they are connected in a different HRR.

@@

    In Figure \ref{fig:connected_percent}, I illustrate the prevalence of overlapping board members or executives over time. The fraction of hospitals with overlapping leaders rises gradually from 10\% in 2014 to 15\% in 2017, then declines slightly after 2017. The peak in 2017 lacks an obvious explanation, but the subsequent decline may reflect growing consolidation in health care, which reduces opportunities to connect with hospitals outside the same system. I also plot the fraction of markets that include at least two hospitals with overlapping leaders. This measure increases from 15\% in 2014 to nearly 30\% in 2017, then falls to roughly 20\% by 2023. For comparison, about 60\% of hospitals in my sample begin as part of a system, and this fraction rises steadily over time. Although informal affiliations through overlapping leaders occur less often than formal system membership, they remain a notable feature of the U.S. hospital landscape. I also present the geographic distribution of these pairs in Figure \ref{fig:connected_maps}. The blue lines represent hospitals that share a common board member or executive, and the red dots are all other hospitals. Connected hospitals are distributed fairly evenly across the US, apart from relatively few connections on the west coast relative to the population.

@@

    \begin{figure}[ht!]
        \centering
        \caption{Geographic Distribution of Board/Executive Interlock}
        \includegraphics[width=.85\textwidth]{Objects/connected_maps.pdf}
        \label{fig:connected_maps}
    \end{figure}

    \subsection{Characteristics of Hospitals with Overlap}

    Hospitals in the sample fall into three categories: always connected, never connected, or connected at some point in time. Table \ref{tab:hosp_group_stats} reports characteristics for each group. Hospitals that never exhibit leadership overlap, whether within or across markets, differ markedly from connected hospitals, with substantially fewer beds and nurses. In contrast, hospitals that become connected during the sample period resemble connected hospitals much more closely. Over 300 hospitals gain overlap within the same HRR, and almost 400 gain overlap across different HRRs. These groups are not mutually exclusive, so some hospitals experience overlap in both settings.\footnote{I account for this in the estimation.} Nearly 600 hospitals remain connected across different HRRs throughout the sample, making this form of overlap the most common.

@@

    I now present characteristics of connected pairs of hospitals in Table \ref{hospital_pair_types}. In the first column, I show pair characteristics of hospitals connected within the same HRR. In the second column, I present characteristics of hospitals connected to otherwise unaffiliated hospitals located in a different HRR. Therefore, all hospital pairs in the table are connected, but only the first column are connected to hospitals that are potentially direct competitors. Each hospital in a pair is general or specialty (I only consider adult hospitals for these categories), children or adult, and independent or belonging to a system, all characteristics in the AHA data. Additionally, the median bed size of all hospitals in the sample is 119, and I define hospitals as being either small or large depending on whether they fall below or above this median. There are typically 1.7 connections in the same HRR and almost 8 connections in different HRRs.

@@

    I will first discuss hospital pairs located in the same HRR. Most of these pairs, 64\%, consist of two general hospitals, and are thus competitors in terms of typical services provided. An additional 12\% of these pairs are between a general and a specialty hospital, and the remaining 1\% are between two specialty hospitals. Similarly, three quarters of the pairs in the same HRR are between two adult hospitals. While the largest fraction of these pairs, 44\%, is between two independent hospitals, there is also a significant portion, 40\%, that occur between one independent hospital and one system-affiliated hospital. Only 16\% of these pairs are between two hospitals belonging to different systems. Almost half of the pairs are between one small and one large hospitals, and the average bed difference is over 200, indicating a significant size difference on average. Finally, pairs within the same HRR are typically 35 miles apart.

    Now, I discuss hospitals that share leaders with hospitals in different markets. 80\% of these pairs are between two general hospitals, and 90\% are between two adult hospitals. Thus, it is likely that hospitals are seeking expertise from other hospitals similar to theirs moreso than in the same HRR. There is a fairly even spread of whether the hospitals in the pair are independent. Still, 50\% of pairs are between one large and one small hospital, and the bed size difference is 238 on average. As expected, the hospitals in these pairs are located much farther apart geographically than those in the same HRR.

    Hospital decisions to become connected are obviously endogenous. Thus, when thinking about how overlap affects hospital behavior, the relevant comparison group is important. To think about this further, I present trends over time for a select number of outcomes for four different groups of hospitals: (1) those that gain overlap in the same HRR, (2) those that gain overlap in a different HRR, (3) those that are always connected in a different HRR, and (4) those that are never connected. I show these trends for outcomes of number of discharges, expenses, and number of services offered by the hospital in Figure \ref{fig:outcome_desc_graph}.

    \begin{figure}[ht!]
        \centering
        \caption{Select Outcomes by Overlap Status}
        \includegraphics[width=.7\textwidth]{Objects/outcome_desc_graph.pdf}
        \label{fig:outcome_desc_graph}
    \end{figure}

  These trends show that not all hospitals serve as a suitable comparison group for those that gain a connection within the same HRR. In the top panel, which plots the average number of services over time, hospitals that remain unconnected or always connected in a different HRR follow a different trajectory than hospitals that become connected. A similar pattern appears for discharges: hospitals that gain overlap track a comparable path before the connection, then diverge, while hospitals that never gain overlap exhibit a distinct trend of declining Medicare and Medicaid discharges. Finally, total operating expenses reveal entirely different slopes for hospitals that never gain overlap compared to those that do, whether in the same or a different market. For this reason, when estimating the effect of gaining overlap in the same market, I compare these hospitals only to those that become connected in a different market. This approach allows me to isolate the effect of connecting with a competitor from the broader effect of becoming connected.
